\sysfull

### 2017-11-15 paper/coalescing.tex
coalescing section revised

Often, energy available at the transiently-powered device is sufficient to run multiple consecutive tasks without page commit at each task boundary. In this case task-based execution model can better amortize the run time overhead of committing updated pages when a task ends. Hence many tasks can be \emph{merged} into a single virtual task---a process we call \emph{coalescing}.

@@

\textbf{Task Coalescing Strategies.} The simplest strategy will coalesce/de-coalesce fixed number of task at every successful/unsuccessful commit operation, where task coalescing factor increases by $C_U$ at each successful virtual task execution and decreases by $C_D$, otherwise. There are naturally infinite possible strategies for coalescing, i.e. infinite amount of $(C_U,C_D)$ pairs. The caveat of predefined $(C_U,C_D)$ pair is inability to adapt to changing energy availability conditions. A \emph{good} coalescing strategy must moderates the risk of coalescing and capitalize on its benefit by dynamically, \emph{adaptively} determining a coalescing factors $(C_U,C_D)$ that combines many tasks without suffering an excess of wasted work. Therefore we propose to consider two fundamental coalescing strategies for \sys, which we will later assess and compare experimentally:

\begin{enumerate}
	\item \textbf{Power Interrupt History-based Coalescing (HDC):} when more tasks are coalesced without power interrupt $C_U$ increases gradually, and $C_D$ increases, as more task are interrupted before committing; and 
	\item \textbf{Weighted Power Interrupt History-based Coalescing (WHDC):} where coalescing strategy follows the one of HDC, while also taking into account the \emph{expected execution time} of each task in the application.
\end{enumerate}

\subsection{Power Interrupt History-driven Coalescing}

With this strategy \sys calculates only one parameter $C=C_U=C_D$ based on the number of tasks completed, until $t$th task commit, $S_t$, during the current operational period. \sys begins execution in a coalesced task with $S_t=0$. When $S_t=C_{t-1}-1$ coalesced tasks have been executed without power interrupt (where $C_{t-1}$ defines $C$ until $t-1$ task commit) a task commit follows. After each commit, \sys assigns new coalescing factor as $C_{t}=\lceil \sum_{k=1}^{t}S_{k-1}/2\rceil$. This implies that if more tasks complete in an operating period, \sys attempts to coalesce more tasks. However, tasks coalescing are increased by a half the speed of the successful task completion. This way we adapt to the coalescing speed/wasted effort trade-off: at the potential power interrupt. In the power failure case, if fewer tasks complete, \sys attempts to coalesce fewer tasks.

@@

\subsection{Weighted Power Interrupt History-driven Coalescing}

@@

\textbf{Task Size Profiling.} To measure how large the distribution of tasks is, we have used a set of program benchmarks (refer to Section~\ref{sec:results_software} for details and to the beginnin of Section~\ref{sec:methodology_evaluation} for system setup). Each benchmark was run for predefined period of time (X\,seconds \todo{Provide exact number}{Amjad}), on non-intermittent power, and at each task evocation, its execution time was measured. If a task was evoked more than once, longer execution time replaced the short one---this way we obtained conservative values of task sizes. At the end of task measurement, normalized values of task execution are assigned for each task, the shortest one being one task, the longest one X tasks \todo{Provide exact number}{Amjad}. Result of the task size measurement is provided in Figure~\ref{fig:task_profiling}. We observe huge variations of task sizes---both within a benchmark, as well per benchmark, which calls for the use of this information in coalescing.

@@

In the second coalescing strategy an application needs to be profiled, as above, before it is loaded to the intermittently-powered device.

>>>>

Often, energy available at the transiently-powered device is sufficient to run multiple consecutive tasks without page commit at each task boundary. This way task-based execution model can better amortize the overhead of committing updated pages when a task ends. Hence many tasks can be \emph{merged} into a single virtual task---a process we call \emph{coalescing}.

@@

\textbf{Task Coalescing Strategies.} The simplest strategy will coalesce/de-coalesce fixed number of task at every successful/unsuccessful commit operation, where task coalescing factor increases by $C_U$ at each successful virtual task execution and decreases by $C_D$, otherwise. There are naturally infinite possible strategies for coalescing, i.e. infinite amount of $(C_U,C_D)$ pairs. The caveat of predefined $(C_U,C_D)$ pair is inability to adapt to changing energy availability conditions. A \emph{good} coalescing strategy must moderate the risk of coalescing and capitalize on its benefit by dynamically, \emph{adaptively} determining a coalescing factors $(C_U,C_D)$ that combines many tasks without suffering an excess of wasted work. Therefore we propose to consider two fundamental coalescing strategies for \sys, which we will later assess and compare experimentally:

\begin{enumerate}
	\item \textbf{Interrupt History Coalescing (HDC):} when more tasks are coalesced without power interrupt, $C_U$ increases, and $C_D$ increases, as more task are interrupted before committing; and 
	\item \textbf{Weighted Interrupt History Coalescing (WHDC):} where coalescing strategy follows the one of HDC, while also taking into account the \emph{expected execution time} of each task in the application.
\end{enumerate}

\subsection{Power Interrupt History Coalescing}

With this strategy \sys calculates only one parameter $C=C_U=C_D$ based on the number of tasks completed, until $t$th task commit, $S_t$, during the current operational period. \sys begins execution in a coalesced task with $S_t=0$. When $S_t=C_{t-1}-1$ coalesced tasks have been executed without power interrupt (where $C_{t-1}$ defines $C$ until $t-1$ task commit) a task commit follows. After each commit, \sys assigns new coalescing factor as $C_{t}=\lceil \sum_{k=1}^{t}S_{k-1}/2\rceil$. This implies that if more tasks complete in an operating period, \sys attempts to coalesce more tasks. However, tasks coalescing are increased by a half the speed of the successful task completion. This way we adapt to the coalescing speed/wasted effort trade-off: at the potential power interrupt \sys does not immediately reduce coalescing factor to one, and with successful page commit of coalescing task, \sys gradually increases coalescing factor. In both cases \sys reduces the effect of outliers, i.e. short bursts of energy or short bursts of frequent power interrupts.

@@

\subsection{Weighted Power Interrupt History Coalescing}

@@

\textbf{Task Size Profiling.} We have used a set of program benchmarks (refer to Section~\ref{sec:results_software} for details and to the beginning of Section~\ref{sec:methodology_evaluation} for system setup). Each benchmark was run for predefined period of time (in case of Figure~\ref{fig:task_profiling} 10\,seconds), on non-intermittent power, and at each task evocation, its execution time was measured. If a task was evoked more than once, longer execution time replaced the short one---this way we obtained conservative values of task sizes. At the end of task measurement, normalized integer values of task execution for task $x$, $W_x$, are collected---the shortest one being one task, the longest one $W_{x,\max}$ (in case of Figure~\ref{fig:task_profiling} $W_{x,\max}=5$). Result of the task size measurement is provided in Figure~\ref{fig:task_profiling}. We observe huge variations of task sizes---both within a benchmark, as well per benchmark, which calls for the use of this information in coalescing.

@@

With the task size information, in the second (weighted) coalescing strategy an application needs to be profiled, as above, before it is loaded to the intermittently-powered device. Then the process of coalescing is the same as for the (non-weighted) power interrupt history coalescing, with the exception that $S_t=W_x \times S_t$, i.e. number of traversed tasks before commit are a function of their size.
